Gender disparity in computational biology research publications

### 2017-08-21 manuscript.tex

% Use the asterisk to denote corresponding authorship and provide email address in note below.
* kevin\_bonham\@hms.harvard.edu

@@

When looking at research output in the form of publication and impact, the story is complex: Women tend to publish less than men \cite{Lariviere2013}, are underrepresented in the more prestigious first and last author positions, and publish fewer single-author papers \cite{West2013}. In mathematics, women tend to publish in lower-impact journals \cite{Mihaljevic-Brandt2016}, while in engineering, women publish in journals with higher impact factors \cite{Ghiasi2015}. In general, however, articles authored by women are cited less frequently than articles authoreed by men \cite{Lariviere2013,Ghiasi2015}, which might in part be due to men citing their own work more often than women do \cite{King2016}. Inferring bias in these studies is difficult, since the cause of the disparity between male and female authorship cannot be readily determined. At the same time, when stories of scientific discoveries are told, gender biases are readily identified: Work by female scientists is more likely to be attributed to a male colleague \cite{Rossiter1993}, and biographies of successful female scientists perpetuate gender stereotypes \cite{Fara2013}. Finally, the way in which evidence for gender bias is received is in itself biased:  Male scientists are less likely to accept studies that point to the existence of gender bias than are their female colleagues \cite{Handley2015}.

>>>>

% Use the asterisk to denote corresponding authorship and provide email address in note below.
* kevbonham\@gmail.com

@@

When looking at research output in the form of publication and impact, the story is complex: Women tend to publish less than men \cite{Lariviere2013}, are underrepresented in the more prestigious first and last author positions, and publish fewer single-author papers \cite{West2013}. In mathematics, women tend to publish in lower-impact journals \cite{Mihaljevic-Brandt2016}, while in engineering, women publish in journals with higher impact factors \cite{Ghiasi2015}. In general, however, articles authored by women are cited less frequently than articles authored by men \cite{Lariviere2013,Ghiasi2015}, which might in part be due to men citing their own work more often than women do \cite{King2016}. Inferring bias in these studies is difficult, since the cause of the disparity between male and female authorship cannot be readily determined. At the same time, when stories of scientific discoveries are told, gender biases are readily identified: Work by female scientists is more likely to be attributed to a male colleague \cite{Rossiter1993}, and biographies of successful female scientists perpetuate gender stereotypes \cite{Fara2013}. Finally, the way in which evidence for gender bias is received is in itself biased:  Male scientists are less likely to accept studies that point to the existence of gender bias than are their female colleagues \cite{Handley2015}.
